Secure Composition of Robust and Optimising Compilers

### 2023-07-05 popl/main.tex
Redo pass over BG

The intuition behind security is that nothing bad happens.
Properties ($\pi$) make this intuition formally precise, however, they rely on a trace model.
{\em Traces} ($\trace$) are (possibly infinite) sequences of {\em events} ($\event$), which are an abstract representation of observable actions.
The empty event ($\emptyevent$) is an uboservable or unimportant action.
Throughout the paper, empty sequences are denoted with $\hole{\cdot}$ and concatenation of sequences $\trace[_{1}],\trace[_{2}]$ with $\trace[_{1}]\cdot\trace[_{2}]$.
This notation also used for prepending events to traces, i.e., $\event\cdot\trace$.

@@

The {\em behavior} of a whole program $\runtimetermvar$ ($\behav{\runtimetermvar}$) is a set containing all traces the program may emit during execution.
%This paper only considers deterministic languages, where the behavior $\behav{\runtimetermvar}$ is a singleton set for any whole $\runtimetermvar$.
Finally, the satisfaction of a class (or lifted property) is defined as follows:

@@

As described earlier, partial programs cannot be executed, because they lack definitions and, thus, get stuck.
{\em Linking} ($\link{\progvar[_{1}]}{\progvar[_{2}]}$) syntactic descriptions $\progvar[_{1}],\progvar[_{2}]$ can give rise to a whole program.
Note that linking is not defined on programs, but merely on their syntactic descriptions, since programs contain dynamic state.
However, for convenience, linking of programs $\runtimetermvar[_{1}],\runtimetermvar[_{2}]$ shall be defined if both $\runtimetermvar[_{1}]$ and $\runtimetermvar[_{2}]$ have an initial dynamic state.
As typically done in a secure compilation context~\cite{abate2019jour}, this paper also assumes that programs are the result of linking a {\em context} ($\ctx$) and a {\em component} ($\comp$), which are just any syntactic description.
The component part of a whole program is the main object of interest and thus a known syntactic description.
In contrast, the context is usually arbitrary and thus has the role of an active {\em attacker} that can interact with the component by means of whatever features the programming language has.
The following definition extends on \Thmref{def:propsat} to cover {\em robustness} and exploits this interaction between the context and the component:

@@

A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
The syntactic description of programs written in $\src{L}$ is denoted as $\src{\progvar}$, while for ones written in $\trg{L}$ this paper writes $\trg{\progvar}$.
It is considered good if this translation is semantics-preserving, i.e., correct.
Correct compilers may still be insecure compilers.
Programs translated by insecure compilers may violate security properties that were assumed to hold, given the presence of a context--level program.
So, ideally, a compiler is secure~\cite{abate2019jour} in the sense of the following definition:

@@

This section introduces a trace model (\Cref{subsec:propdefs}) that is used to define properties (\gls{tms}, \gls{sms}, \gls{ms}, and \gls{scct}).
The properties are of practical importance and also of interest in the case study (\Cref{sec:casestud}) this paper presents.
The study uses monitors, which are able to check wether a trace admits a property or not.
The use of monitors is because they have some structure useful for inductive proofs, whereas traces lack this structure, thus making proofs difficult.
So, this section also presents monitors for each of the defined properties (\Cref{subsec:monitors}) and introduces monitor-based satisfaction.
Finally, for each property, it is shown that monitor-based satisfaction implies property satisfaction.

>>>>

The intuition behind security is that nothing bad happens.
Properties ($\pi$) make this intuition formally precise, however, they rely on a trace model.
{\em Traces} ($\trace$) are (possibly infinite) sequences of {\em events} ($\event$), which are an abstract representation of observable actions.
The empty event ($\emptyevent$) is an uboservable or unimportant action.
Throughout the paper, empty sequences are denoted with $\hole{\cdot}$ and concatenation of sequences $\trace[_{1}],\trace[_{2}]$ with $\trace[_{1}]\cdot\trace[_{2}]$.
For prepending events to traces the paper uses the same notation, i.e., $\event\cdot\trace$.

@@

The {\em behavior} of a whole program $\runtimetermvar$ ($\behav{\runtimetermvar}$) is a set containing all traces the program may emit during execution.
%This paper only considers deterministic languages, where the behavior $\behav{\runtimetermvar}$ is a singleton set for any whole $\runtimetermvar$.
This set is used to define the satisfaction of a property as follows:

@@

As described earlier, partial programs cannot be executed, because they lack definitions and, thus, get stuck.
{\em Linking} ($\link{\progvar[_{1}]}{\progvar[_{2}]}$) syntactic descriptions $\progvar[_{1}],\progvar[_{2}]$ can give rise to a whole program.
Note that linking is not defined on programs, but merely on their syntactic descriptions, since programs contain dynamic state.
However, for convenience, linking of programs $\runtimetermvar[_{1}],\runtimetermvar[_{2}]$ shall be defined if both $\runtimetermvar[_{1}]$ and $\runtimetermvar[_{2}]$ have an initial dynamic state.
As typically done in a secure compilation context~\cite{abate2019jour}, this paper also assumes that programs are the result of linking a {\em context} ($\ctx$) and a {\em component} ($\comp$), which are just any syntactic description.
The component part of a whole program is the main object of interest and thus a known syntactic description.
In contrast, the context is usually arbitrary and thus has the role of an active {\em attacker} that can interact with the component by means of whatever features the programming language has.
The following extends on \Thmref{def:propsat} to cover {\em robustness} and exploits this interaction between the context and the component:

@@

A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
For the syntactic description of programs written in $\src{L}$, this paper writes $\src{\progvar}$, while for ones written in $\trg{L}$, this paper writes $\trg{\progvar}$.
It is considered good if this translation is semantics-preserving, i.e., correct.
Correct compilers may still be insecure compilers\MK{citation needed}.
Programs translated by insecure compilers may violate security properties that were assumed to hold, given the presence of a context--level program.
So, ideally, a compiler is secure~\cite{abate2019jour} in the sense of the following definition:

@@

This section introduces a trace model (\Cref{subsec:propdefs}) that is used to define properties (\gls{tms}, \gls{sms}, \gls{ms}, and \gls{scct}).
The properties are of practical importance and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) this paper presents.
The study uses monitors, which are able to check wether a trace admits a property or not.
The use of monitors is because they have some structure useful for inductive proofs, whereas traces lack this structure, thus making proofs difficult.
So, this section also presents monitors for each of the defined properties (\Cref{subsec:monitors}) and introduces monitor-based satisfaction.
Finally, for each property, it is shown that monitor-based satisfaction implies property satisfaction.
